---
title: "R Notebook"
output: html_notebook
---

This code imports, analyzes, and plots data from a multiple replcicates of beta-galactosidase assays. This analysis is for experiments examining beta-galactosidase activity as a function of RpoS level.

Before running this code, you must have followed the directions and run the code in "analysis of single beta-gal experiment.Rmd" on every file you intend analyze here. 

```{r, message=FALSE}
library(readxl)
library(tidyverse)
source("betaGalAnalysisFunctions.R")
theme_set(theme_classic())

fileNames <- c("6-6-18 LB32_LB73.xlsx","6-6-18 LB32_LB73 fake data for testing.xlsx")

processedFiles <- map(fileNames, meanExpressionOneReplicate)

combinedData <- bind_rows(processedFiles)
```

```{r}
ggplot(combinedData) + geom_point(aes(x=RpoS, y=meanExpression)) + facet_grid(.~strain) 
```

Plot mean ± SE for each RpoS level
```{r}
dayMeans <- combinedData %>% 
              group_by(strain, RpoS) %>% 
              summarise(meanAcrossDays = mean(meanExpression), sdAcrossDays = sd(meanExpression)/sqrt(length(meanExpression))) %>%
              mutate(upperErrorBar = meanAcrossDays + sdAcrossDays, lowerErrorBar = meanAcrossDays - sdAcrossDays)

ggplot(dayMeans) + 
  geom_point(aes(x = RpoS, y = meanAcrossDays)) + 
  geom_errorbar(aes(x = RpoS, y = meanAcrossDays, ymax = upperErrorBar, ymin = lowerErrorBar)) + 
  facet_grid(.~strain)

  
```
##Sensitivity

This section is specific for analyzing the sensitivity of a promoter. It requires three RpoS levels.

This function expects three RpoS levels, so for this example I'll subset out some, since we have five.

```{r}
dataForSensitivity <- combinedData %>% 
                        filter(RpoS == 0 | RpoS == 19 | RpoS == 94) 
```


Calculate sensitivity of a promoter. 
```{r}
sensitivities <- dataForSensitivity %>% 
                  group_by(strain, date, experimenter) %>%
                  summarise(sensitivity = calculateSingleSens(RpoS, meanExpression))
```


Plot the data with a line to illustrate null expectation used to assess sensitivity. 

If you designed your experiment with only three RpoS levels, delete the the `filter` step. Note that we have to go back a step to use use the `dayMeans` tibble.
```{r}
dayMeans %>% 
  filter(RpoS == 0 | RpoS == 19 | RpoS == 94) %>% 
  ggplot() + 
  geom_point(aes(x = RpoS, y = meanAcrossDays)) + 
  geom_errorbar(aes(x = RpoS, y = meanAcrossDays, ymax = upperErrorBar, ymin = lowerErrorBar)) + 
  facet_grid(.~strain)
                 
```


